A Loop Quantm Gravity Inspired Action for the Bosonic String and Emergent Dimensions at Large Scales

### 2025-07-02 AoP/response-to-referee.tex

\emph{This approach differs from previous minimal length proposals in string theory \citep{Amati1989Canb,Chang2011Minimal,Gross1987Highenergy,Gross1988String,Yoneya1989Interpretation,Yoneya2000String} in that it specifically targets area rather than length, reflecting LQG's emphasis on area as the fundamental geometric observable.}

\subsection{Section 2 - Part (a): Action Justification}

\RC The proposed action $S_{MNG}=-T\int dx^2\sqrt{-(h+k \Delta)}$ needs a clearer justification: the author states in a footnote that $\Delta$ is as a scalar density later identified with the Kalb-Ramond field should be, at least, physically justified because the remaining of this section is based in this argument, otherwise it should be rigorously derived.

\AR We must be honest and state that at present the area modified action is not rigorously derived. We have added a paragraph right after the action area-modified action (eqn 2.2. in the text) to explain this point:

\emph{At present we do not have a rigorous derivation of this expression. It is the simplest possible ansatz one can think of to encode the insight from LQG that at the Planck scale there is a minimum quantum of area. What makes us more confident in this ansatz is that it is consistent with the Kalb-Ramond field identification and the minimal area interpretation of the scalar density $\Delta$. However, this was not known prior to coming up with the ansatz and only discovered after the fact.}

@@

\subsection{Section 2 - Part (b): Expansion Parameter}

\RC The expansion in $k=l_{pl}/l_s$ seems sound, but the assumption $k\ll 1$ (Planck scale much smaller than string scale) should be discussed further—does this hold in all regimes?

\AR We've added the following paragraph right after the expansion parameter $k$ definition to explain our choice of $k \ll 1$:

\emph{If we take $k$ to be an expansion parameter, then it follows that it should also be small. The physical justification for this is the following picture. The microscopic picture of geometry at the Planck scale, which we adopt as the ``correct'' picture, is that given by loop quantum gravity in terms of spin networks and spin foams. The worldsheet picture emerges when one considers a large enough section of the spin network with a sufficiently large number of punctures. In this limit, the spin network can be approximated by a smooth geometry, and the worldsheet picture emerges as a low-energy effective description of the spin network. When $k \sim 1$, the discrete structure of the underlying spin network becomes manifest and the smooth worldsheet approximation is no longer valid, indicating that a full quantum gravitational treatment - i.e. where geometry itself is fully quantum - is required.}

@@

\subsection{Section 2 - Part (c): Duality Interpretation}

\RC The duality between $h_{ab}$ and $h^{ab}$ (Eq. 2.6a-2.6c) is intriguing but lacks a deeper physical interpretation. How does this duality manifest in observable string dynamics?

\AR We have added the following paragraph which hopefully addresses this point:

\emph{This duality is reminiscent of, but also distinct from, the T and S dualities in string theory. In T-duality the behavior of a string compactified on a circle of radius $R$ is related to the behavior of a string compactified on a circle of radius $\alpha'/R$. In S-duality the physics at strong string coupling $g_s \gg 1$ can be mapped to the physics at weak string coupling $g_s \ll 1$. In our case, the duality is between a description of physics at large string tension $T$ and small worldsheet areas $h_{ab}$ and a description of physics at small string tension $T'$ and large worldsheet areas $h'_{ab} = (h_{ab})^{-1}$. This appears to have features of both T-duality and S-duality, in that the area inversion is reminiscent of T-duality and the string tension inversion is reminiscent of S-duality's strong-weak coupling correspondence.}

### 2025-07-02 lqg-strings.tex
T2: Address referee's Section 2 concerns: action justification, parameter validity, duality interpretation
- Added acknowledgment of ansatz nature in manuscript after S_MNG equation
- Justified k<<1 expansion using LQG spin network emergence picture
- Connected h_ab/h^ab duality to T/S-duality framework as novel feature
- Completed reviewer responses for Section 2(a), 2(b), 2(c) with physical justifications
- Added Luigi's bachelor thesis files
- Updated memory bank

In a very real sense string theory is a quantum theory of geometry. After all, the Nambu-Goto action:
\begin{equation}\label{eqn:ng-action}
	S_{NG} = -T \int \diff ^2 x \sqrt{- \det h}
\end{equation}
is nothing more than the \emph{area} of the string worldsheet. So when we are constructing the quantum theory of the string we are constructing a theory in which the fundamental excitations are geometric in nature. What we would like to do is to modify the string action in a way which incorporates the insight from LQG that at the Planck scale there is a minimum quantum of area. One simple way in which to do this is to modify the action \eqref{eqn:ng-action} as follows:
\begin{equation} \label{eqn:ng-area-corrected}
	S_{M N G} = - T \int \diff ^2 x \sqrt{- (h + k \Delta)}.
\end{equation}
Here, $k = l_{pl} / l_s$, where $l_{pl}$ is the Planck scale, the scale at which quantum geometric effects become important; $l_s$ is the string scale where quantum geometric effects due to area quantization become small; for the time being we take $\Delta$ to be a scalar constant which corresponds to the minimum quantum of area at the Planck scale\footnote{We will see later that $\Delta$ is, in fact, a scalar density of weight 2 which can be taken to be the determinant of an antisymmetric worldsheet tensor $\Delta _{a b} (x)$ and corresponds to the square of the quantum of area at the point $x$ where the edges of the bulk spin network state pierce the WS and endow it with quanta of area as described in the previous section.}. In the limit that $l_s \gg l_{pl}$, $k \ll 1$ and the string action reduces to the conventional Nambu-Goto action with small corrections coming from quantum geometry. In this limit we can expand the integrand as follows:
\begin{equation} \label{eqn:inverse-metric-correction}
	\sqrt{- (h + k \Delta)} \sim \sqrt{- h}\left(1 - \frac{1}{2} \frac{k \Delta}{(- h)} + \mc{O}(k^2/h^2)\right).
\end{equation}
Presumably the limit $k \ll 1$ corresponds to when the area of the string worldsheet is large compared to $\Delta$, $h \gg \Delta$. Thus in this limit $ k^2 \ll 1$ and $ h^2 \gg 1$ because of which we can drop the $\mc{O}(k^2/h^2)$ corrections. The final area corrected Nambu-Goto action becomes:
\begin{equation} \label{eqn:modified-ng-action}
	S'_{NG} = -T \int \diff ^2 x \left( \sqrt{- h} - \frac{1}{2} \frac{k \Delta}{\sqrt{- h}} \right).
\end{equation}
This action has, to the best of our knowledge, not been explored in the string theory literature. However, such an inverse determinant term arises very naturally in the LQG approach and goes by the name of ``inverse triad'' corrections, in the context of the 3D volume of spatial hypersurfaces.

The action eq \eqref{eqn:modified-ng-action} has manifest background Poincaré invariance since just like the regular Nambu-Goto action since all its terms are derivatives of $X^{\mu}$ with all Lorentz indices contracted, thus are Lorentz scalars. At first glance it might seem to not be WS reparameterization invariant because of the second term, however, this is corrected by the fact that $\Delta$ is a scalar \emph{density} of weight 2, thus counteracting the inverse metric determinant. Moreover, if we look at that expression we see that it can be written as:
\begin{equation} \label{eqn:modified-ng-action-v2}
	S'_{NG}[h_{ab}] = -T \int \diff ^2 x \left( \sqrt{- h} - \frac{1}{2} {k \Delta}{\sqrt{- (h^{-1})}} \right),
\end{equation}
where $h^{-1}$ is now the determinant of the \emph{inverse} metric $h^{ab}$. As such, in the expression \eqref{eqn:modified-ng-action} we can see a duality between the first and second terms. If we replace the metric $h_{ab}$ by its inverse $h'_{ab} \equiv h^{ab}$, then the action in terms of $h'_{ab}$ is the same as \eqref{eqn:modified-ng-action}, but with the difference that the factor of $k \Delta / 2$ is mapped to $2 / (k \Delta)$:
\begin{align} \label{eqn:modified-ng-action-v3}
	S'_{NG}[h'_{ab}] & = -T' \int \diff ^2 x' \left( \sqrt{- h'} - \frac{2}{k \Delta} \sqrt{- (h')^{-1}} \right) \nonumber \\
	                 & = -T' \int \diff ^2 x' \left( \sqrt{- h'} - \onehalf {k' \Delta'}\sqrt{- (h')^{-1}} \right),
\end{align}
which has the same form as the original action \eqref{eqn:modified-ng-action}, as long as we make the replacements
\begin{subequations}
	\begin{equation} \label{eqn:T'}
		T \rightarrow T' = - \frac{1}{2} k T
	\end{equation}
	\begin{equation} \label{eqn:k-Delta}
		\frac{1}{2} {k \Delta} \rarrow \frac{1}{2} {k' \Delta'} = \frac{2} {k \Delta}
	\end{equation}
	\begin{equation} \label{eqn:d2x}
		\diff ^2 x \rightarrow \diff ^2 x' = \Delta \diff ^2 x
	\end{equation}
\end{subequations}
(or equivalently $\diff x^{a} \rightarrow \diff x' _{a} = \Delta _{a b} \diff x^{b}$). This is a duality between the physics at the Planck scale and the physics at the string scale, \ie between strong coupling ($k' \gg 1$) and weak coupling ($k \ll 1$), small quantum of area ($\Delta$) and large quantum of area ($\Delta'$) and between large string tension $T$ and small string tension $T'$.

>>>>

In a very real sense string theory is a quantum theory of geometry. After all, the Nambu-Goto action:
\begin{equation}\label{eqn:ng-action}
	S_{NG} = -T \int \diff ^2 x \sqrt{- \det h}
\end{equation}
is nothing more than the \emph{area} of the string worldsheet. So when we are constructing the quantum theory of the string we are constructing a theory in which the fundamental excitations are geometric in nature. What we would like to do is to modify the string action in a way which incorporates the insight from LQG that at the Planck scale there is a minimum quantum of area. One simple way in which to do this is to modify the action \eqref{eqn:ng-action} as follows:
\begin{equation} \label{eqn:ng-area-corrected}
	S_{M N G} = - T \int \diff ^2 x \sqrt{- (h + k \Delta)}.
\end{equation}
At present we do not have a rigorous derivation of this expression. It is the simplest possible ansatz one can think of to encode the insight from LQG that at the Planck scale there is a minimum quantum of area. What makes us more confident in this ansatz is that it is consistent with the Kalb-Ramond field identification and the minimal area interpretation of the scalar density $\Delta$. However, this was not known prior to coming up with the ansatz and only discovered after the fact.

Here, $k = l_{pl} / l_s$, where $l_{pl}$ is the Planck scale, the scale at which quantum geometric effects become important; $l_s$ is the string scale where quantum geometric effects due to area quantization become small; for the time being we take $\Delta$ to be a scalar constant which corresponds to the minimum quantum of area at the Planck scale\footnote{We will see later that $\Delta$ is, in fact, a scalar density of weight 2 which can be taken to be the determinant of an antisymmetric worldsheet tensor $\Delta _{a b} (x)$ and corresponds to the square of the quantum of area at the point $x$ where the edges of the bulk spin network state pierce the WS and endow it with quanta of area as described in the previous section.}. In the limit that $l_s \gg l_{pl}$, $k \ll 1$ and the string action reduces to the conventional Nambu-Goto action with small corrections coming from quantum geometry. In this limit we can expand the integrand as follows:
\begin{equation} \label{eqn:inverse-metric-correction}
	\sqrt{- (h + k \Delta)} \sim \sqrt{- h}\left(1 - \frac{1}{2} \frac{k \Delta}{(- h)} + \mc{O}(k^2/h^2)\right).
\end{equation}

If we take $k$ to be an expansion parameter, then it follows that it should also be small. The physical justification for this is the following picture. The microscopic picture of geometry at the Planck scale, which we adopt as the ``correct'' picture, is that given by loop quantum gravity in terms of spin networks and spin foams. The worldsheet picture emerges when one considers a large enough section of the spin network with a sufficiently large number of punctures. In this limit, the spin network can be approximated by a smooth geometry, and the worldsheet picture emerges as a low-energy effective description of the spin network. When $k \sim 1$, the discrete structure of the underlying spin network becomes manifest and the smooth worldsheet approximation is no longer valid, indicating that a full quantum gravitational treatment - i.e. where geometry itself is fully quantum - is required.

The limit $k \ll 1$ corresponds to when the area of the string worldsheet is large compared to $\Delta$, $h \gg \Delta$. Thus in this limit $ k^2 \ll 1$ and $ h^2 \gg 1$ because of which we can drop the $\mc{O}(k^2/h^2)$ corrections. The final area corrected Nambu-Goto action becomes:
\begin{equation} \label{eqn:modified-ng-action}
	S'_{NG} = -T \int \diff ^2 x \left( \sqrt{- h} - \frac{1}{2} \frac{k \Delta}{\sqrt{- h}} \right).
\end{equation}
This action has, to the best of our knowledge, not been explored in the string theory literature. However, such an inverse determinant term arises very naturally in the LQG approach and goes by the name of ``inverse triad'' corrections, in the context of the 3D volume of spatial hypersurfaces.

The action eq \eqref{eqn:modified-ng-action} has manifest background Poincaré invariance since just like the regular Nambu-Goto action since all its terms are derivatives of $X^{\mu}$ with all Lorentz indices contracted, thus are Lorentz scalars. At first glance it might seem to not be WS reparameterization invariant because of the second term, however, this is corrected by the fact that $\Delta$ is a scalar \emph{density} of weight 2, thus counteracting the inverse metric determinant. Moreover, if we look at that expression we see that it can be written as:
\begin{equation} \label{eqn:modified-ng-action-v2}
	S'_{NG}[h_{ab}] = -T \int \diff ^2 x \left( \sqrt{- h} - \frac{1}{2} {k \Delta}{\sqrt{- (h^{-1})}} \right),
\end{equation}
where $h^{-1}$ is now the determinant of the \emph{inverse} metric $h^{ab}$. As such, in the expression \eqref{eqn:modified-ng-action} we can see a duality between the first and second terms. If we replace the metric $h_{ab}$ by its inverse $h'_{ab} \equiv h^{ab}$, then the action in terms of $h'_{ab}$ is the same as \eqref{eqn:modified-ng-action}, but with the difference that the factor of $k \Delta / 2$ is mapped to $2 / (k \Delta)$:
\begin{align} \label{eqn:modified-ng-action-v3}
	S'_{NG}[h'_{ab}] & = -T' \int \diff ^2 x' \left( \sqrt{- h'} - \frac{2}{k \Delta} \sqrt{- (h')^{-1}} \right) \nonumber \\
	                 & = -T' \int \diff ^2 x' \left( \sqrt{- h'} - \onehalf {k' \Delta'}\sqrt{- (h')^{-1}} \right),
\end{align}
which has the same form as the original action \eqref{eqn:modified-ng-action}, as long as we make the replacements
\begin{subequations}
	\begin{equation} \label{eqn:T'}
		T \rightarrow T' = - \frac{1}{2} k T
	\end{equation}
	\begin{equation} \label{eqn:k-Delta}
		\frac{1}{2} {k \Delta} \rarrow \frac{1}{2} {k' \Delta'} = \frac{2} {k \Delta}
	\end{equation}
	\begin{equation} \label{eqn:d2x}
		\diff ^2 x \rightarrow \diff ^2 x' = \Delta \diff ^2 x
	\end{equation}
\end{subequations}
(or equivalently $\diff x^{a} \rightarrow \diff x' _{a} = \Delta _{a b} \diff x^{b}$). This is a duality between the physics at the Planck scale and the physics at the string scale, \ie between strong coupling ($k' \gg 1$) and weak coupling ($k \ll 1$), small worldsheet areas $h_{ab}$ and large worldsheet areas $h'_{ab}$ and between large string tension $T$ and small string tension $T'$.

This duality is reminiscent of, but also distinct from, the T and S dualities in string theory. In T-duality the behavior of a string compactified on a circle of radius $R$ is related to the behavior of a string compactified on a circle of radius $\alpha'/R$. In S-duality the physics at strong string coupling $g_s \gg 1$ can be mapped to the physics at weak string coupling $g_s \ll 1$. In our case, the duality is between a description of physics at large string tension $T$ and small worldsheet areas $h_{ab}$ and a description of physics at small string tension $T'$ and large worldsheet areas $h'_{ab} = (h_{ab})^{-1}$. This appears to have features of both T-duality and S-duality, in that the area inversion is reminiscent of T-duality and the string tension inversion is reminiscent of S-duality's strong-weak coupling correspondence.
